High Contrast Coronagraphy

### 2025-01-10 src/tex/ms.tex
RvC colour and typos

\usepackage{hyperref}
\usepackage[version=4]{mhchem}
\usepackage{booktabs}% http://ctan.org/pkg/booktabs
\newcommand{\tabitem}{~~\llap{\textbullet}~~}
\hypersetup{colorlinks=true,allcolors=[rgb]{0,0,0.8}}

@@

\newcommand{\project}[1]{\textsf{#1}}

% Document starts
\begin{document}

@@

\section{INTRODUCTION}
\label{sec:intro}

@@

Initially developed to image the Sun's corona without the need for a Solar eclipse \citep{Lyot33}, one of the most significant science drivers for the latest coronagraphs is in the detection and characterisation of circumstellar material and planets around nearby stars.
%
Young self-luminous gas giant exoplanets have been discovered around young stars \citep[see ][ for a review of these detections]{Zurlo24} both in the nearby Galactic field and further away in young stellar OB associations (out to 400 pc), typically these exoplanets have luminosities of $10^{-4}-10^{-6}$ of their parent star at angular separations up to a few arcseconds.

The search for life beyond the Earth has focused on the idea that water is an essential part of life cycles elsewhere in the Universe, as it is a polar solvent formed from two elements that are found in abundance throughout the Galaxy.
%
Places where water can exist in its liquid state form prime locations for these searches, notably Earth-like planets and ice moons that have a liquid water ocean underneath an ice layer.
%
The region around a star where liquid water can exist on the surface of a terrestrial planet (with appropriate atmospheric pressure) is referred to as the \ac{hz}; for the Sun this is from 0.9-1.2 au but this can move as the luminosity of stars change with time.

At optical wavelengths, exoplanets are not self-luminous but instead reflect the light of their parent star.
%
For an Earth analogue with similar radius, albedo and effective temperature orbiting around a solar-type star 10 parsecs away, the typical amount of reflected light in the optical is $10^{-10}$ of the central star at a separation of 0.1 arcseconds at maximum elongation.
%
The technical challenge is in distinguishing the light of the parent star from the light of the planet.
%
Planetary systems that are closer to the Sun have two benefits: one that is twice as close as the other will have double the angular separation between the star and planet, and from the inverse square law, four times more flux is received from the planet.
%
For direct imaging, therefore, the closest stars to the Sun are the ones that are studied for the presence of directly imaged exoplanets.
%
On average there are more M dwarf stars close to the Sun than other solar type stars: in a volume limited (20 parsec) sample around the Sun, there are $\sim 140$ solar G-type stars, and on the order of 1900 M-dwarf stars \citep{Kirkpatrick24}.
%
For solar type stars, the contrast required in the optical wavelengths is on the order of $10^{-10}$ for terrestrial planets in the \ac{hz}, but for smaller mass stars with lower luminosities, the contrast is $10^{-7}$ for M dwarfs - for stars around the Sun this is shown in Figure~\ref{fig:nstars}.

@@

The markers for biosignatures also set the parameters (such as wavelength and bandwidth) that form part of the design decision.
%
Spectra of Earthshine (light reflected from the Earth by the non-Sun lit portion of the Moon) from 0.5 to 2.4 microns show \ce{O2}, \ce{O3}, \ce{CO2}, \ce{CH4} and \ce{H2O} \citep{Turnbull06}, and there are many discussions about the possible biomarker molecules that should be searched for and characterised as evidence for biosignatures \citep[see the reviews of ][]{2016AsBio..16..465S,2017ARAA..55..433K,2018AsBio..18..663S}.
%
It is clear that the unambiguous detection of life will require not one singular detection of a molecule, but will be the combination of several different lines of evidence.
%
We focus on the technical challenges for imaging a rocky terrestrial planet around a nearby star in our Galaxy.

@@

Here the $z$ component of the wave vector is defined as $k_z=\sqrt{n^2k^2 - k_{\perp}^2}$ and $\mathcal{F}_{x,y}^{(-1)}$ is defined as the (inverse) Fourier transform over the x and y coordinates.
%
While Equation \ref{eq:angular_spectrum} describes the full propagation from one plane to another, it is quite unwieldy to use and does not provide much physical insight.
%
For many optical systems it is sufficient to analyze the paraxial performance.
%
The paraxial approximation assumes that the plane waves make small angles with respect to the z-axis which means that the $x$ and $y$ wave vector components are $k_x,k_y<<1$.
%
In this regime the propagation factor $k_z=\sqrt{n^2k^2 - k_{\perp}^2}\approx nk - \frac{k_{\perp}^2}{2nk}$ and the Fresnel propagation equation becomes
\begin{equation}
\label{eq:angular_spectrum2}
\mathcal{E}(x, y, z') = e^{-ink(z'-z)} \mathcal{F}_{x,y}^{-1}\{ e^{-i\frac{k_{\perp}^2}{2nk}} \mathcal{F}_{x,y}\{\mathcal{E}(x,y,z)\}\}.
\end{equation}
The real space solution that is derived from the propagation equation is the classical Fresnel diffraction equation,
\begin{equation}
\label{eq:fresnel_diffraction}
\mathcal{E}(x, y, z') = \frac{e^{-ink(z'-z)}}{\lambda z} \iint_{-\infty}^{\infty} \mathcal{E}(u, v, z) e^{\frac{ik\left[(x - u)^2 + (y - v)^2\right]}{2 (z' - z)}} \mathrm{d}u \mathrm{d}v.
\end{equation}
The far-field approximation ($z \gg x, y$) and the substitution $\theta_x = \frac{x}{\lambda z}$, $\theta_y = \frac{y}{\lambda z}$ results in the Fraunhofer diffraction integral,
\begin{equation}
\label{eq:fraunhofer_diffraction}
\mathcal{E}(\theta_x, 
\theta_y) \propto \iint_{-\infty}^{\infty} \mathcal{E}(u, v) e^{-i2\pi(\theta_x u + \theta_y v)} \mathrm{d}u \mathrm{d}v.
\end{equation}
This shows that the far-field distribution of an electromagnetic wave is its Fourier transform.
%
A perfectly aligned lens, or the mirrors of a telescope, brings rays that come from infinity to the focal point.
%
This is identical to a far-field transform, and therefore the operation of a lens can be described by the Fraunhofer diffraction integral.

@@

Apodizing the telescope pupil provides an opportunity to redistribute the light in the \ac{scfp} to form `dark zones' around the target star where an exoplanet can be imaged.
%
Strictly this is not so much a coronagraph as a modification to the \ac{psf} of the instrument - all objects in the focal plane have the same \ac{psf}, both stars and exoplanets together.
%
As long as the angular diameter of the star is smaller than the Airy core of the \ac{psf}, \ac{pp} coronagraphs are not impacted by the diameter of the star or by residual tip tilt vibrations that are not removed by the \ac{ao} control loop, making them a robust alternative to the more efficient, smaller \ac{iwa} \ac{fp} coronagraphs. 
%
Suppressing diffraction in the \ac{psf} requires destructive interference using coherent light from the Airy core, but at a cost in Strehl ratio of the \ac{psf}.
%
Since the exoplanet \ac{psf} is identical to the stellar \ac{psf}, the planet throughput decreases too.

@@

The first \ac{app} optics were chromatic: the OPD was a function of the refractive index of the transmissive material, with the suppression decreasing with increasing bandwidth.
%
Achromatic phase shifts can be implemented using the principle of geometric phase: the vector-APP \citep[vAPP; ][]{Snik12} replaces the classical phase pattern $\phi_{\textrm{c}}(\vec{r}) = n(\lambda) \Delta d(\vec{r})$,
with the ``geometric phase'' \citep[known as the Pancharatnam-Berry phase; ][]{Pancharatnam,Berry}.
%
The vAPP phase pattern is imposed by a half-wave retarder with a patterned fast axis orientation $\theta(\vec{r})$.
%
The geometric phase is imprinted on incident beams decomposed according to circular polarization state: $\phi_{\textrm{g}}(\vec{r}) = \pm2\cdot\theta(\vec{r})$, with the sign depending on the circular polarization handedness.
% %%%
As this fast axis orientation pattern does not vary as a function of wavelength (with the possible exception of an inconsequential offset/piston term), the geometric phase is strictly achromatic.
%
Vector-APP devices are produced by applying two liquid-crystal techniques: any desired phase pattern is applied onto a substrate glass through a \textit{direct-write procedure} \citep{directwrite} that applies the orientation pattern $\theta(\vec{r})$ by locally polymerizing the alignment layer material in the direction set by the controllable polarization of a scanning UV laser.
%
Consecutive layers of birefringent liquid-crystal are deposited on top of this alignment layer, which subsequently self-align \citep[``\textit{Multi-Twist Retarders}''; MTR ][]{MTR} with predetermined parameters (birefringence dispersion, thickness, nematic twist) to yield a linear retardance that is close to half-wave over the specified wavelength range.
%
Additional layers broaden the wavelength range to over an octave in wavelength, at a cost of an absorption feature due to the carbon-carbon bonds within the liquid crystal and glue layers.
%
The vector-APP devices required additional optics (typically a half-wave plate and Wollaston prism) to isolate the two circular polarizations and produce two separate \acp{psf} with dark holes on opposing sides of the central star \citep{Snik12}.
%
By adding a phase diffraction grating onto the APP phase pattern to make a grating vector APP \citep[gvAPP; ][]{Snik12,Otten14}, the other optics are no longer required.
%
The two coronagraphic \acp{psf} are separated diffracted into the $m=\pm 1$ order, with a $m=0$ ``leakage term'' non-coronagraphic \ac{psf} with flux of a few percent of the original star left in the undeviated beam, acting both as an astrometric and photometric reference \citep{Otten17,Sutlieff24}.
%
The grating effect means that the \ac{psf} centroids vary as a function of wavelength, and so the gvAPPs are ideal for imaging onto integral field units and image slicers \citep{Sutlieff21,Sutlieff23}.
%
This liquid-crystal technology has enabled coronagraphic designs that were previously impossible to manufacture, including the coronagraphic modal \ac{wfs} \citep{Wilby17}, sparse aperture masking with multiple holograms \citep{Doelman21}, complex amplitude Vector Vortex Coronagraphs \citep[VVC; ][]{Snik14}, and triple grating coronagraphs \citep{Doelman20} that redisperse the \acp{psf} back into white light coronagraphic \acp{psf} for VVCs \citep{Doelman23,Laginga24}. 
%
A comprehensive review of the coronagraphs enabled by the liquid crystal technology is given in \citet{Doelman2021a}.

@@

Original \ac{piaa} designs are for unobscured circular apertures: telescope pupils with secondary obscurations require reformatting of the pupil to make a continuous diffraction-free \ac{psf} in the coronagraphic focal plane.
%
The introduction of complex masks that can be manufactured to the required tolerances enable \ac{piaa}s for complex and segmented telescope pupils, suitable for space-based telescope designs such as the HabEx/LUVOIR concepts.
%
Results from the laboratory demonstration of a \ac{piaacmc} with a segmented aperture, \citet{Marx21} show contrasts of $3\times 10^{-8}$ from $4-9$\ld{} with 2\% bandwidth.
%
Most recently, the laboratory demonstration of high contrast with the \ac{piaacmc} coronagraph on an obstructed and segmented aperture \citep{Belikov22} shows $1.9\times 10^{-8}$ contrast achieved in a 10\% bandwidth between $3.5-8$\ld{}.

Ultimately the rejected light can form the basis for a \ac{wfs} to keep the \ac{piaa} pointed and aligned with the science target, and an integrated \ac{wfs} and coronagraph with \ac{piaacmc} has been demonstrated \citep{Haffert23a}.

>>>>

\usepackage{hyperref}
\usepackage[version=4]{mhchem}
\usepackage{booktabs}% http://ctan.org/pkg/booktabs
\newcommand{\tabitem}{~~\llap{\textbullet}~~}
%\hypersetup{colorlinks=true}

@@

\newcommand{\project}[1]{\textsf{#1}}
\hypersetup{%
        colorlinks,
        breaklinks=true,
        plainpages=false,%
        citecolor=[rgb]{0,0.20,0.45},
        linkcolor=[rgb]{0,0.20,0.45},
        urlcolor=[rgb]{0,0.20,0.45},
        bookmarksopen=true,%
        bookmarksnumbered=false,%
        bookmarksdepth=5%
}
% Document starts
\begin{document}

@@

\section{Introduction}
\label{sec:intro}

@@

Initially developed to image the Sun's corona without the need for a Solar eclipse \citep{Lyot33}, one of the most significant science drivers for the latest coronagraphs is in the detection and characterisation of circumstellar material and planets around nearby stars.
%
Young self-luminous gas giant exoplanets have been directly imaged at infrared wavelengths around young stars \citep[see ][ for a review of these detections]{Zurlo24} both in the nearby Galactic field and further away in young stellar OB associations (out to 400 pc), typically these exoplanets have luminosities of $10^{-4}-10^{-6}$ of their parent star at angular separations up to a few arcseconds.

The search for life beyond the Earth has focused on the idea that water is an essential part of life cycles elsewhere in the Universe, as it is a polar solvent formed from two elements that are found in abundance throughout the Galaxy.
%
Places where water can exist in its liquid state form prime locations for these searches, notably Earth-like planets and ice moons that have a liquid water ocean underneath an ice layer.
%
The region around a star where liquid water can exist on the surface of a terrestrial planet (with appropriate atmospheric pressure) is referred to as the \ac{hz}; for the Sun this is from 0.9-1.2 au but this can move as the luminosity of stars evolve over time.

At visible light wavelengths, the flux from exoplanets that can be directly imaged is overwhelmingly dominated by the reflected light from their parent star.
%
For an Earth analogue with similar radius, albedo and effective temperature orbiting around a solar-type star 10 parsecs away, the typical amount of reflected light in the optical is $10^{-10}$ of the central star at a separation of 0.1 arcseconds at maximum elongation.
%
The technical challenge is in distinguishing the light of the parent star from the light of the planet.
%
Planetary systems that are closer to the Sun have two benefits: one that is twice as close as the other will have double the angular separation between the star and planet, and from the inverse square law, four times more flux is received from the planet.
%
For direct imaging, therefore, the closest stars to the Sun are the ones that are studied for the presence of directly imaged exoplanets.
%
On average there are more M dwarf stars close to the Sun than other solar type stars: in a volume limited (20 parsec) sample around the Sun, there are $\sim 140$ solar G-type stars, and on the order of 1900 M-dwarf stars \citep{Kirkpatrick24}.
%
For solar type stars, the contrast required in the optical wavelengths is on the order of $10^{-10}$ for terrestrial planets in the \ac{hz}, but for smaller mass stars with lower luminosities, the contrast is $10^{-7}$ for M dwarfs - for stars around the Sun this is shown in Figure~\ref{fig:nstars}.

@@

The markers for biosignatures also set the parameters (such as wavelength and bandwidth) that form part of the design decision.
%
Spectra of Earthshine (sunlight that is reflected off the Earth and illuminates the dark portion of the Moon) from 0.5 to 2.4 microns show \ce{O2}, \ce{O3}, \ce{CO2}, \ce{CH4} and \ce{H2O} \citep{Turnbull06}, and there are many discussions about the possible biomarker molecules that should be searched for and characterised as evidence for biosignatures \citep[see the reviews of ][]{2016AsBio..16..465S,2017ARAA..55..433K,2018AsBio..18..663S}.
%
It is clear that the unambiguous detection of life will require not one singular detection of a molecule, but will be the combination of several different lines of evidence.
%
We focus on the technical challenges for imaging a rocky terrestrial planet around a nearby star in our Galaxy.

@@

Here the $z$ component of the wave vector is defined as $k_z=\sqrt{n^2k^2 - k_{\perp}^2}$ and $\mathcal{F}_{x,y}^{(-1)}$ is defined as the (inverse) Fourier transform over the x and y coordinates.
%
While Equation \ref{eq:angular_spectrum} describes the full propagation from one plane to another, it is quite unwieldy to use and does not provide much physical insight.
%
For many optical systems it is sufficient to analyze the paraxial performance.
%
The paraxial approximation assumes that the plane waves make small angles with respect to the z-axis which means that the $x$ and $y$ wave vector components are $k_x,k_y<<1$.
%
In this regime the propagation factor $k_z=\sqrt{n^2k^2 - k_{\perp}^2}\approx nk - \frac{k_{\perp}^2}{2nk}$ simplifying to:
\begin{equation}
\label{eq:angular_spectrum2}
\mathcal{E}(x, y, z') = e^{-ink(z'-z)} \mathcal{F}_{x,y}^{-1}\{ e^{-i\frac{k_{\perp}^2}{2nk}} \mathcal{F}_{x,y}\{\mathcal{E}(x,y,z)\}\}.
\end{equation}
The real space solution that is derived from this propagation equation is the classical Fresnel diffraction equation,
\begin{equation}
\label{eq:fresnel_diffraction}
\mathcal{E}(x, y, z') = \frac{e^{-ink(z'-z)}}{\lambda z} \iint_{-\infty}^{\infty} \mathcal{E}(u, v, z) e^{\frac{ik\left[(x - u)^2 + (y - v)^2\right]}{2 (z' - z)}} \mathrm{d}u \mathrm{d}v.
\end{equation}
The far-field approximation ($z \gg x, y$) and the substitution $\theta_x = \frac{x}{\lambda z}$, $\theta_y = \frac{y}{\lambda z}$ results in the Fraunhofer diffraction integral,
\begin{equation}
\label{eq:fraunhofer_diffraction}
\mathcal{E}(\theta_x, 
\theta_y) \propto \iint_{-\infty}^{\infty} \mathcal{E}(u, v) e^{-i2\pi(\theta_x u + \theta_y v)} \mathrm{d}u \mathrm{d}v.
\end{equation}
This shows that the far-field distribution of an electromagnetic wave is its Fourier transform.
%
A perfectly aligned lens, or the mirrors of a telescope, brings rays that come from infinity to the focal point.
%
This is identical to a far-field transform, and therefore the operation of a lens can be described by the Fraunhofer diffraction integral.

@@

Apodizing the telescope pupil provides an opportunity to redistribute the light in the \ac{scfp} to form `dark zones' around the target star where an exoplanet can be imaged.
%
Strictly this is not so much a coronagraph as a modification to the \ac{psf} of the instrument - all objects in the focal plane have the same \ac{psf}, both stars and exoplanets together.
%
As long as the angular diameter of the star is smaller than the Airy core of the \ac{psf}, \ac{pp} coronagraphs are not impacted by the diameter of the star or by residual tip tilt vibrations that are not removed by the \ac{ao} control loop, making them a robust alternative to the more efficient, smaller \ac{iwa} \ac{fp} coronagraphs. 
%
Suppressing diffraction in the \ac{psf} requires destructive interference using coherent light from the Airy core, but decreases the encircled energy of the core of the \ac{psf}.
%
Since the exoplanet \ac{psf} is identical to the stellar \ac{psf}, the planet throughput decreases too.

@@

The first \ac{app} optics were chromatic: the Optical Path Difference (OPD) was a function of the refractive index of the transmissive material, with the suppression decreasing with increasing bandwidth.
%
Achromatic phase shifts can be implemented using the principle of geometric phase: the vector-APP \citep[vAPP; ][]{Snik12} replaces the classical phase pattern $\phi_{\textrm{c}}(\vec{r}) = n(\lambda) \Delta d(\vec{r})$,
with the ``geometric phase'' \citep[known as the Pancharatnam-Berry phase; ][]{Pancharatnam,Berry}.
%
The vAPP phase pattern is imposed by a half-wave retarder with a patterned fast axis orientation $\theta(\vec{r})$.
%
The geometric phase is imprinted on incident beams decomposed according to circular polarization state: $\phi_{\textrm{g}}(\vec{r}) = \pm2\cdot\theta(\vec{r})$, with the sign depending on the circular polarization handedness.
% %%%
As this fast axis orientation pattern does not vary as a function of wavelength (with the possible exception of an inconsequential offset/piston term), the geometric phase is strictly achromatic.
%
Vector-APP devices are produced by applying two liquid-crystal techniques: any desired phase pattern is applied onto a substrate glass through a \textit{direct-write procedure} \citep{directwrite} that applies the orientation pattern $\theta(\vec{r})$ by locally polymerizing the alignment layer material in the direction set by the controllable polarization of a scanning UV laser.
%
Consecutive layers of birefringent liquid-crystal are deposited on top of this alignment layer, which subsequently self-align \citep[``\textit{Multi-Twist Retarders}''; MTR ][]{MTR} with predetermined parameters (birefringence dispersion, thickness, nematic twist) to yield a linear retardance that is close to half-wave over the specified wavelength range.
%
Additional layers broaden the wavelength range to over an octave in wavelength, at a cost of an absorption feature due to the carbon-carbon bonds within the liquid crystal and glue layers.
%
The vector-APP devices required additional optics (typically a half-wave plate and Wollaston prism) to isolate the two circular polarizations and produce two separate \acp{psf} with dark holes on opposing sides of the central star \citep{Snik12}.
%
By adding a phase diffraction grating onto the APP phase pattern to make a grating vector APP \citep[gvAPP; ][]{Snik12,Otten14}, the other optics are no longer required.
%
The two coronagraphic \acp{psf} are separated diffracted into the $m=\pm 1$ order, with a $m=0$ ``leakage term'' non-coronagraphic \ac{psf} with flux of a few percent of the original star left in the undeviated beam, acting both as an astrometric and photometric reference \citep{Otten17,Sutlieff24}.
%
The grating effect means that the \ac{psf} centroids vary as a function of wavelength, and so the gvAPPs are ideal for imaging onto integral field units and image slicers \citep{Sutlieff21,Sutlieff23}.
%
This liquid-crystal technology has enabled coronagraphic designs that were previously impossible to manufacture, including the coronagraphic modal \ac{wfs} \citep{Wilby17}, sparse aperture masking with multiple holograms \citep{Doelman21}, complex amplitude Vector Vortex Coronagraphs \citep[VVC; ][]{Snik14}, and triple grating coronagraphs \citep{Doelman20} that redisperse the \acp{psf} back into white light coronagraphic \acp{psf} for VVCs \citep{Doelman23,Laginga24}. 
%
A comprehensive review of the coronagraphs enabled by the liquid crystal technology is given in \citet{Doelman2021a}.

@@

Original \ac{piaa} designs are for unobscured circular apertures: telescope pupils with secondary obscurations require reformatting of the pupil to make a continuous diffraction-free \ac{psf} in the coronagraphic focal plane.
%
The introduction of complex masks that can be manufactured to the required tolerances enable \ac{piaa}s for complex and segmented telescope pupils, suitable for space-based telescope designs such as the HabEx/LUVOIR concepts.
%
Results from the laboratory demonstration of a \ac{piaacmc} with a segmented aperture, \citet{Marx21} show contrasts of $3\times 10^{-8}$ from $4-9$\ld{} with 2\% bandwidth.
%
Most recently, the laboratory demonstration of high contrast with the \ac{piaacmc} coronagraph on an obstructed and segmented aperture \citep{Belikov22} shows $1.9\times 10^{-8}$ contrast achieved in a 10\% bandwidth between $3.5-8$\ld{}.
%
Ultimately the rejected light can form the basis for a \ac{wfs} to keep the \ac{piaa} pointed and aligned with the science target, and an integrated \ac{wfs} and coronagraph with \ac{piaacmc} has been demonstrated \citep{Haffert23a}.
